---
title: "ESM 204 Assignment 4"
author: "Madeline Oliver"
date: "5/19/2020"
output: html_document
---

```{r setup, echo = FALSE, message = FALSE, warning = FALSE, cache = FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, cache = FALSE)

library(tidyverse)
library(here)
library(janitor)
library(lubridate)
library(chron)
library(stringr)
library(pracma)
```

Equations:
temp = temperature in year t
econ = the fraction of economic activity that is retained in a year if the temp is "temp"
consumption = total consumption at time t
utility = society's utility from consumption
Ramsey rule:
  o = pure rate of time preference
  g = growth rate of income over time
  n = elasticity of marginal utility of consumption with respect to income (how much more does a poor person value a dollar compared to a rich person?)

```{r}
# Assign values
T_low <- 0
T_med <- 2
T_high <- 8
T_bau <- 4.4
t <- 0:200
o <- 0.005
n <- .5
g <- .01
b <- .05
```

```{r}
# Build functions in this chunk
temp <- function(t, T) {temp = pmin(T*t/100, T)}
econ <- function(t) {econ = exp(-1*b*t^2)}
consumption <- function(k, t) {k*exp(g*t)}
utility <- function(c) {(c^(1-n))/(1-n)}
discount_rate <- o + n*g
```

```{r}
# Build data frames for each scenario
T_00 <- data.frame(temp_increase = T_low,
                   year = t,
                   temp = temp(t, T_low)) %>% 
  mutate(economy = econ(temp)) %>% 
  mutate(consumption = consumption(economy, year)) %>% 
  mutate(utility = utility(consumption)) %>% 
  mutate(pv_utility = utility/(1+discount_rate)^year)

T_02 <- data.frame(temp_increase = T_med,
                   year = t,
                   temp = temp(t, T_med)) %>% 
  mutate(economy = econ(temp)) %>% 
  mutate(consumption = consumption(economy, year)) %>% 
  mutate(utility = utility(consumption))

T_08 <- data.frame(temp_increase = T_high,
                   year = t,
                   temp = temp(t, T_high)) %>% 
  mutate(economy = econ(temp)) %>% 
  mutate(consumption = consumption(economy, year)) %>% 
  mutate(utility = utility(consumption)) 

T_04.4 <- data.frame(temp_increase = T_bau,
                   year = t,
                   temp = temp(t, T_bau)) %>% 
  mutate(economy = econ(temp)) %>% 
  mutate(consumption = consumption(economy, year)) %>% 
  mutate(utility = utility(consumption)) %>% 
  mutate(pv_utility = utility/(1+discount_rate)^year) %>% 
  data.frame(T_00$pv_utility) %>% 
  mutate(pct_change = (pv_utility - T_00.pv_utility)/T_00.pv_utility*100)

L <- data.frame(sum(T_04.4$T_00.pv_utility),
                           sum(T_04.4$pv_utility)) %>%
  rename(pv_utility_base = sum.T_04.4.T_00.pv_utility.,
         pv_utility_bau = sum.T_04.4.pv_utility.) %>%
  mutate(L = (pv_utility_bau - pv_utility_base)/pv_utility_base)

```

```{r}
# Plot everything using data frames created above
# Temperature
ggplot()+
  geom_line(data = T_00, aes(x = year, y = temp, color = "No Climate Change"), size = 1)+
  geom_line(data = T_02, aes(x = year, y = temp, color  = "Modest Climate Change"), size = 1)+
  geom_line(data = T_08, aes(x = year, y = temp, color = "Extreme Climate Change"), size = 1) +
  labs(title = "Temperature over time under various climate scenarios",
       color = "Climate Scenario") +
  scale_color_manual(values = c("orangered4", "orangered", "orange")) +
  theme(plot.title = element_text(
    family = "Times",
    face = "bold",
    size = 12),
    legend.text = element_text(family = "Times"),
    legend.title = element_text(family = "Times"))

# Consumption
ggplot()+
  geom_line(data = T_00, aes(x = year, y = consumption, color = "No Climate Change"), 
            size = 1)+
  geom_line(data = T_02, aes(x = year, y = consumption, color = "Modest Climate Change"), 
            size = 1)+
  geom_line(data = T_08, aes(x = year, y = consumption, color = "Extreme Climate Change"), 
            size = 1) +
  labs(title = "Consumption over time under various climate scenarios",
       color = "Climate Scenario") +
  scale_color_manual(values = c("orangered4", "orangered", "orange")) +
  theme(plot.title = element_text(
    family = "Times",
    face = "bold",
    size = 12),
    legend.text = element_text(family = "Times"),
    legend.title = element_text(family = "Times"))
  
# Utility
ggplot()+
  geom_line(data = T_00, aes(x = year, y = utility,  color = "No Climate Change"), size = 1)+
  geom_line(data = T_02, aes(x = year, y = utility, color = "Modest Climate Change"), size = 1)+
  geom_line(data = T_08, aes(x = year, y = utility, color = "Extreme Climate Change"), size = 1) +
  labs(title = "Undiscounted utility over time under various climate scenarios",
       color = "Climate Scenario") +
  scale_color_manual(values = c("orangered4", "orangered", "orange")) +
  theme(plot.title = element_text(
    family = "Times",
    face = "bold",
    size = 12),
    legend.text = element_text(family = "Times"),
    legend.title = element_text(family = "Times"))
```
Formula for present value:
net utility today = SUM[utility at year t / (1+discount_rate)^t] across all years


Part 2.B.

10% increase in $\delta$
```{r}
# Assign values
T_low <- 0
T_med <- 2
T_high <- 8
T_bau <- 4.4
t <- 0:200
o <- 0.005*1.1
n <- .5
g <- .01
b <- .05
# Build functions in this chunk
temp <- function(t, T) {temp = pmin(T*t/100, T)}
econ <- function(t) {econ = exp(-1*b*t^2)}
consumption <- function(k, t) {k*exp(g*t)}
utility <- function(c) {(c^(1-n))/(1-n)}
discount_rate <- o + n*g
# Build data frames for each scenario
T_00 <- data.frame(temp_increase = T_low,
                   year = t,
                   temp = temp(t, T_low)) %>% 
  mutate(economy = econ(temp)) %>% 
  mutate(consumption = consumption(economy, year)) %>% 
  mutate(utility = utility(consumption)) %>% 
  mutate(pv_utility = utility/(1+discount_rate)^year)

T_04.4 <- data.frame(temp_increase = T_bau,
                   year = t,
                   temp = temp(t, T_bau)) %>% 
  mutate(economy = econ(temp)) %>% 
  mutate(consumption = consumption(economy, year)) %>% 
  mutate(utility = utility(consumption)) %>% 
  mutate(pv_utility = utility/(1+discount_rate)^year) %>% 
  data.frame(T_00$pv_utility) %>% 
  mutate(pct_change = (pv_utility - T_00.pv_utility)/T_00.pv_utility*100)

L_o <- data.frame(sum(T_04.4$T_00.pv_utility),
                           sum(T_04.4$pv_utility)) %>%
  rename(pv_utility_base = sum.T_04.4.T_00.pv_utility.,
         pv_utility_bau = sum.T_04.4.pv_utility.) %>%
  mutate(L_o = (pv_utility_bau - pv_utility_base)/pv_utility_base) %>%
  mutate(L = L$L) %>%
  mutate(elasticity = .1/((L_o-L)/L)) %>%
  mutate(pct_change_L = elasticity*.1)
```

The delta elasticity of utility is `r L_o$elasticity`
The pct change in L is `r L_o$pct_change_L` 

10% increase in $\eta$
```{r}
# Assign values
T_low <- 0
T_med <- 2
T_high <- 8
T_bau <- 4.4
t <- 0:200
o <- 0.005
n <- .5*1.1
g <- .01
b <- .05
# Build functions in this chunk
temp <- function(t, T) {temp = pmin(T*t/100, T)}
econ <- function(t) {econ = exp(-1*b*t^2)}
consumption <- function(k, t) {k*exp(g*t)}
utility <- function(c) {(c^(1-n))/(1-n)}
discount_rate <- o + n*g
# Build data frames for each scenario
T_00 <- data.frame(temp_increase = T_low,
                   year = t,
                   temp = temp(t, T_low)) %>% 
  mutate(economy = econ(temp)) %>% 
  mutate(consumption = consumption(economy, year)) %>% 
  mutate(utility = utility(consumption)) %>% 
  mutate(pv_utility = utility/(1+discount_rate)^year)

T_04.4 <- data.frame(temp_increase = T_bau,
                   year = t,
                   temp = temp(t, T_bau)) %>% 
  mutate(economy = econ(temp)) %>% 
  mutate(consumption = consumption(economy, year)) %>% 
  mutate(utility = utility(consumption)) %>% 
  mutate(pv_utility = utility/(1+discount_rate)^year) %>% 
  data.frame(T_00$pv_utility) %>% 
  mutate(pct_change = (pv_utility - T_00.pv_utility)/T_00.pv_utility*100)

L_n <- data.frame(sum(T_04.4$T_00.pv_utility),
                           sum(T_04.4$pv_utility)) %>%
  rename(pv_utility_base = sum.T_04.4.T_00.pv_utility.,
         pv_utility_bau = sum.T_04.4.pv_utility.) %>%
  mutate(L_n = (pv_utility_bau - pv_utility_base)/pv_utility_base) %>%
  mutate(L = L$L) %>%
  mutate(elasticity = .1/((L_n-L)/L))%>%
  mutate(pct_change_L = elasticity*.1)
```

The eta elasticity of utility is `r L_n$elasticity`
The pct change in L is `r L_n$pct_change_L` 


10% increase in g
```{r}
# Assign values
T_low <- 0
T_med <- 2
T_high <- 8
T_bau <- 4.4
t <- 0:200
o <- 0.005
n <- .5
g <- .01*1.1
b <- .05
# Build functions in this chunk
temp <- function(t, T) {temp = pmin(T*t/100, T)}
econ <- function(t) {econ = exp(-1*b*t^2)}
consumption <- function(k, t) {k*exp(g*t)}
utility <- function(c) {(c^(1-n))/(1-n)}
discount_rate <- o + n*g
# Build data frames for each scenario
T_00 <- data.frame(temp_increase = T_low,
                   year = t,
                   temp = temp(t, T_low)) %>% 
  mutate(economy = econ(temp)) %>% 
  mutate(consumption = consumption(economy, year)) %>% 
  mutate(utility = utility(consumption)) %>% 
  mutate(pv_utility = utility/(1+discount_rate)^year)

T_04.4 <- data.frame(temp_increase = T_bau,
                   year = t,
                   temp = temp(t, T_bau)) %>% 
  mutate(economy = econ(temp)) %>% 
  mutate(consumption = consumption(economy, year)) %>% 
  mutate(utility = utility(consumption)) %>% 
  mutate(pv_utility = utility/(1+discount_rate)^year) %>% 
  data.frame(T_00$pv_utility) %>% 
  mutate(pct_change = (pv_utility - T_00.pv_utility)/T_00.pv_utility*100)

L_g <- data.frame(sum(T_04.4$T_00.pv_utility),
                           sum(T_04.4$pv_utility)) %>%
  rename(pv_utility_base = sum.T_04.4.T_00.pv_utility.,
         pv_utility_bau = sum.T_04.4.pv_utility.) %>%
  mutate(L_g = (pv_utility_bau - pv_utility_base)/pv_utility_base) %>%
  mutate(L = L$L) %>%
  mutate(elasticity = .1/((L_g-L)/L))%>%
  mutate(pct_change_L = elasticity*.1)
```

The gamma elasticity of utility is `r L_g$elasticity`
The pct change in L is `r L_g$pct_change_L` 


10% increase in $\beta$
```{r}
# Assign values
T_low <- 0
T_med <- 2
T_high <- 8
T_bau <- 4.4
t <- 0:200
o <- 0.005
n <- .5
g <- .01
b <- .05*1.1
# Build functions in this chunk
temp <- function(t, T) {temp = pmin(T*t/100, T)}
econ <- function(t) {econ = exp(-1*b*t^2)}
consumption <- function(k, t) {k*exp(g*t)}
utility <- function(c) {(c^(1-n))/(1-n)}
discount_rate <- o + n*g
# Build data frames for each scenario
T_00 <- data.frame(temp_increase = T_low,
                   year = t,
                   temp = temp(t, T_low)) %>% 
  mutate(economy = econ(temp)) %>% 
  mutate(consumption = consumption(economy, year)) %>% 
  mutate(utility = utility(consumption)) %>% 
  mutate(pv_utility = utility/(1+discount_rate)^year)

T_04.4 <- data.frame(temp_increase = T_bau,
                   year = t,
                   temp = temp(t, T_bau)) %>% 
  mutate(economy = econ(temp)) %>% 
  mutate(consumption = consumption(economy, year)) %>% 
  mutate(utility = utility(consumption)) %>% 
  mutate(pv_utility = utility/(1+discount_rate)^year) %>% 
  data.frame(T_00$pv_utility) %>% 
  mutate(pct_change = (pv_utility - T_00.pv_utility)/T_00.pv_utility*100)

L_b <- data.frame(sum(T_04.4$T_00.pv_utility),
                           sum(T_04.4$pv_utility)) %>%
  rename(pv_utility_base = sum.T_04.4.T_00.pv_utility.,
         pv_utility_bau = sum.T_04.4.pv_utility.) %>%
  mutate(L_b = (pv_utility_bau - pv_utility_base)/pv_utility_base) %>%
  mutate(L = L$L) %>%
  mutate(elasticity = .1/((L_b-L)/L)) %>%
  mutate(pct_change_L = elasticity*.1)
```

The beta elasticity of utility is `r L_b$elasticity`
The pct change in L is `r L_b$pct_change_L` 

10% increase in T_bau
```{r}
# Assign values
T_low <- 0
T_med <- 2
T_high <- 8
T_bau <- 4.4*1.1
t <- 0:200
o <- 0.005
n <- .5
g <- .01
b <- .05
# Build functions in this chunk
temp <- function(t, T) {temp = pmin(T*t/100, T)}
econ <- function(t) {econ = exp(-1*b*t^2)}
consumption <- function(k, t) {k*exp(g*t)}
utility <- function(c) {(c^(1-n))/(1-n)}
discount_rate <- o + n*g
# Build data frames for each scenario
T_00 <- data.frame(temp_increase = T_low,
                   year = t,
                   temp = temp(t, T_low)) %>% 
  mutate(economy = econ(temp)) %>% 
  mutate(consumption = consumption(economy, year)) %>% 
  mutate(utility = utility(consumption)) %>% 
  mutate(pv_utility = utility/(1+discount_rate)^year)

T_04.4 <- data.frame(temp_increase = T_bau,
                   year = t,
                   temp = temp(t, T_bau)) %>% 
  mutate(economy = econ(temp)) %>% 
  mutate(consumption = consumption(economy, year)) %>% 
  mutate(utility = utility(consumption)) %>% 
  mutate(pv_utility = utility/(1+discount_rate)^year) %>% 
  data.frame(T_00$pv_utility) %>% 
  mutate(pct_change = (pv_utility - T_00.pv_utility)/T_00.pv_utility*100)

L_t <- data.frame(sum(T_04.4$T_00.pv_utility),
                           sum(T_04.4$pv_utility)) %>%
  rename(pv_utility_base = sum.T_04.4.T_00.pv_utility.,
         pv_utility_bau = sum.T_04.4.pv_utility.) %>%
  mutate(L_t = (pv_utility_bau - pv_utility_base)/pv_utility_base) %>%
  mutate(L = L$L) %>%
  mutate(elasticity = .1/((L_t-L)/L))%>%
  mutate(pct_change_L = elasticity*.1)
```

The temperature elasticity of utility is `r L_t$elasticity`
The pct change in L is `r L_t$pct_change_L` 

2.c. 
$\theta$-

```{r}
# Assign values, back to the OG parameters
T_low <- 0
T_med <- 2
T_high <- 8
T_bau <- 4.4
t <- 0:200
o <- 0.005
n <- .5
g <- .01
b <- .05
```

```{r}
# Build functions in this chunk
temp <- function(t, T) {temp = pmin(T*t/100, T)}
econ <- function(t) {econ = exp(-1*b*t^2)}
consumption <- function(k, t) {k*exp(g*t)}
consumption_o <- function(k, t, theta) {k*(exp(g*t))*theta}
utility <- function(c) {(c^(1-n))/(1-n)}
discount_rate <- o + n*g
theta <-  function(c4, c0, n) {((c4^(1-n))/c0)^(1/(1-n))}
```

```{r}
# Build data frames for each scenario
theta_01 <- fzero(fun = function(theta){((theta)^(1-n)*L$pv_utility_base) - L$pv_utility_bau}, x = 0)$x

theta <- 1-theta_01
discount_rate <- o + n*g
consumption <- function(k, t) {theta*(k*exp(g*t))}


# Create data frame at T = 0 using the new consumption function and theta*
# ------------------------------------------------------------------------
T_00 <- data.frame(temp_increase = T_low,
                   year = t,
                   temp = temp(t, T_low)) %>% 
  mutate(economy = econ(temp)) %>% 
  mutate(consumption = consumption(economy, year)) %>% 
  mutate(utility = utility(consumption)) %>% 
  mutate(pv_utility = utility/(1+discount_rate)^year)
# Sum the NPV of utility at T = 0 with theta* and compare with NPV of utility at T = 4.4 to double check
# ---------------------------------------------------------------------------------------------------------
theta_01 <- data.frame(sum(T_00$pv_utility),
                       L$pv_utility_bau) %>% 
  rename(utility_base = sum.T_00.pv_utility.) %>% 
  rename(utility_bau = L.pv_utility_bau)

```

When we are certain that under business as usual, T = 4.4, society is willing to give up `r theta` of their consumption every year for the next 200 years in order to completely prevent climate change. 


```{r}
# # Question 2d - Goal: Identify the value of theta (theta*) at which the present value of utility at T = 0 is equal to the present value of utiity at e(T) ~ 4.2. First, calculate the expected value of T. Assign theta a value of 1. 
# # -----------------------------------------------------------------------
# T_exp <- 2*.2 + 4*.5 + 6*.3
# theta <- 1
# # Create data frame for T = e(T) ~ 4.2
# # -------------------------------------
# T_exp_01 <- data.frame(temp_increase = T_exp,
#                    year = t,
#                    temp = temp(t, T_exp)) %>% 
#   mutate(economy = econ(temp)) %>% 
#   mutate(consumption = consumption(economy, year)) %>% 
#   mutate(utility = utility(consumption)) %>% 
#   mutate(pv_utility = utility/(1+discount_rate)^year)
# # Sum the pv of utility at e(T)
# # -----------------------------
# T_exp_sum <- data.frame(sum(T_exp_01$pv_utility)) %>% 
#   rename(utility_exp = sum.T_exp_01.pv_utility.)
```

```{r}
theta <- 1
T_exp <- 2
# Create data frame for T = e(T) ~ 4.2
# -------------------------------------
T_exp_2 <- data.frame(temp_increase = T_exp,
                   year = t,
                   temp = temp(t, T_exp)) %>% 
  mutate(economy = econ(temp)) %>% 
  mutate(consumption = consumption(economy, year)) %>% 
  mutate(utility = utility(consumption)) %>% 
  mutate(pv_utility = utility/(1+discount_rate)^year)

T_exp <- 4
T_exp_4 <- data.frame(temp_increase = T_exp,
                   year = t,
                   temp = temp(t, T_exp)) %>% 
  mutate(economy = econ(temp)) %>% 
  mutate(consumption = consumption(economy, year)) %>% 
  mutate(utility = utility(consumption)) %>% 
  mutate(pv_utility = utility/(1+discount_rate)^year)

T_exp <- 6
T_exp_6 <- data.frame(temp_increase = T_exp,
                   year = t,
                   temp = temp(t, T_exp)) %>% 
  mutate(economy = econ(temp)) %>% 
  mutate(consumption = consumption(economy, year)) %>% 
  mutate(utility = utility(consumption)) %>% 
  mutate(pv_utility = utility/(1+discount_rate)^year)

T_exp_sum <- data.frame(sum(T_exp_2$pv_utility)) %>% 
  rename(utility_exp_2 = sum.T_exp_2.pv_utility.) %>%
  data.frame(sum(T_exp_4$pv_utility)) %>% 
  rename(utility_exp_4 = sum.T_exp_4.pv_utility.) %>%
  data.frame(sum(T_exp_6$pv_utility)) %>% 
  rename(utility_exp_6 = sum.T_exp_6.pv_utility.) %>%
  mutate(utility_exp = utility_exp_2*.2 + utility_exp_4*.5 + utility_exp_6*.3)

```


```{r}
# Question 2d continued
# ---------------------
# Anthony's solution to 2c 
# --------------------------------------------------------------------------------------
theta_star_02 <- fzero(fun = function(theta) {((theta)^(1-n)*L$pv_utility_base) - T_exp_sum$utility_exp}, 
                       x = 0)$x
# Assign new value to theta
# -------------------------
theta <- 1-theta_star_02
# Create data frame for T = 0 using the new consumption function and theta*
# -------------------------------------------------------------------------
T_00 <- data.frame(temp_increase = T_low,
                   year = t,
                   temp = temp(t, T_low)) %>% 
  mutate(economy = econ(temp)) %>% 
  mutate(consumption = consumption(economy, year)) %>% 
  mutate(utility = utility(consumption)) %>% 
  mutate(pv_utility = utility/(1+discount_rate)^year)
# Sum the PV of utility at T = 0 with theta* and compare with NPV of utility at T = 4.4 to ensure accuracy
# ---------------------------------------------------------------------------------------------------------
theta_02 <- data.frame(sum(T_00$pv_utility),
                       T_exp_sum$utility_exp) %>% 
  rename(utility_base = sum.T_00.pv_utility.) %>% 
  rename(utility_bau = T_exp_sum.utility_exp)
```

Under uncertainty over T, society is willing to give up `r theta` of their consumption every year for the next 200 years in order to completely prevent climate change. 



