---
title: "ESM 204 Assignment 4"
author: "Madeline Oliver"
date: "5/19/2020"
output: html_document
---

```{r setup, echo = FALSE, message = FALSE, warning = FALSE, cache = FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, cache = FALSE)

library(tidyverse)
library(here)
library(janitor)
library(lubridate)
library(chron)
library(stringr)
```

Equations:
temp = temperature in year t
econ = the fraction of economic activity that is retained in a year if the temp is "temp"
consumption = total consumption at time t
utility = society's utility from consumption
Ramsey rule:
  o = pure rate of time preference
  g = growth rate of income over time
  n = elasticity of marginal utility of consumption with respect to income (how much more does a poor person value a dollar compared to a rich person?)

```{r}
# Assign values
T_low <- 0
T_med <- 2
T_high <- 8
T_bau <- 4.4
t <- 0:200
o <- 0.005
n <- .5
g <- .01
b <- .05
```

```{r}
# Build functions in this chunk
temp <- function(t, T) {temp = pmin(T*t/100, T)}
econ <- function(t) {econ = exp(-1*b*t^2)}
consumption <- function(k, t) {k*exp(g*t)}
utility <- function(c) {(c^(1-n))/(1-n)}
discount_rate <- o + n*g
```

```{r}
# Build data frames for each scenario
T_00 <- data.frame(temp_increase = T_low,
                   year = t,
                   temp = temp(t, T_low)) %>% 
  mutate(economy = econ(temp)) %>% 
  mutate(consumption = consumption(economy, year)) %>% 
  mutate(utility = utility(consumption)) %>% 
  mutate(pv_utility = utility/(1+discount_rate)^year)

T_02 <- data.frame(temp_increase = T_med,
                   year = t,
                   temp = temp(t, T_med)) %>% 
  mutate(economy = econ(temp)) %>% 
  mutate(consumption = consumption(economy, year)) %>% 
  mutate(utility = utility(consumption))

T_08 <- data.frame(temp_increase = T_high,
                   year = t,
                   temp = temp(t, T_high)) %>% 
  mutate(economy = econ(temp)) %>% 
  mutate(consumption = consumption(economy, year)) %>% 
  mutate(utility = utility(consumption)) 

T_04.4 <- data.frame(temp_increase = T_bau,
                   year = t,
                   temp = temp(t, T_bau)) %>% 
  mutate(economy = econ(temp)) %>% 
  mutate(consumption = consumption(economy, year)) %>% 
  mutate(utility = utility(consumption)) %>% 
  mutate(pv_utility = utility/(1+discount_rate)^year) %>% 
  data.frame(T_00$pv_utility) %>% 
  mutate(pct_change = (pv_utility - T_00.pv_utility)/T_00.pv_utility*100)

L <- data.frame(sum(T_04.4$T_00.pv_utility),
                           sum(T_04.4$pv_utility)) %>%
  rename(pv_utility_base = sum.T_04.4.T_00.pv_utility.,
         pv_utility_bau = sum.T_04.4.pv_utility.) %>%
  mutate(L = (pv_utility_bau - pv_utility_base)/pv_utility_base)

```

```{r}
# Plot everything using data frames created above
# Temperature
ggplot()+
  geom_line(data = T_00, aes(x = year, y = temp, color = "No Climate Change"), size = 1)+
  geom_line(data = T_02, aes(x = year, y = temp, color  = "Modest Climate Change"), size = 1)+
  geom_line(data = T_08, aes(x = year, y = temp, color = "Extreme Climate Change"), size = 1) +
  labs(title = "Temperature over time under various climate scenarios",
       color = "Climate Scenario") +
  scale_color_manual(values = c("orangered4", "orangered", "orange")) +
  theme(plot.title = element_text(
    family = "Times",
    face = "bold",
    size = 12))

# Consumption
ggplot()+
  geom_line(data = T_00, aes(x = year, y = consumption, color = "No Climate Change"), 
            size = 1)+
  geom_line(data = T_02, aes(x = year, y = consumption, color = "Modest Climate Change"), 
            size = 1)+
  geom_line(data = T_08, aes(x = year, y = consumption, color = "Extreme Climate Change"), 
            size = 1) +
  labs(title = "Consumption over time under various climate scenarios",
       color = "Climate Scenario") +
  scale_color_manual(values = c("orangered4", "orangered", "orange")) +
  theme(plot.title = element_text(
    family = "Times",
    face = "bold",
    size = 12))
  
# Utility
ggplot()+
  geom_line(data = T_00, aes(x = year, y = utility,  color = "No Climate Change"), size = 1)+
  geom_line(data = T_02, aes(x = year, y = utility, color = "Modest Climate Change"), size = 1)+
  geom_line(data = T_08, aes(x = year, y = utility, color = "Extreme Climate Change"), size = 1) +
  labs(title = "Undiscounted utility over time under various climate scenarios",
       color = "Climate Scenario") +
  scale_color_manual(values = c("orangered4", "orangered", "orange")) +
  theme(plot.title = element_text(
    family = "Times",
    face = "bold",
    size = 12))
```


Formula for present value:
net utility today = SUM[utility at year t / (1+discount_rate)^t] across all years



Part 2.B.

10% increase o
```{r}
# Assign values
T_low <- 0
T_med <- 2
T_high <- 8
T_bau <- 4.4
t <- 0:200
o <- 0.005*1.1
n <- .5
g <- .01
b <- .05
# Build functions in this chunk
temp <- function(t, T) {temp = pmin(T*t/100, T)}
econ <- function(t) {econ = exp(-1*b*t^2)}
consumption <- function(k, t) {k*exp(g*t)}
utility <- function(c) {(c^(1-n))/(1-n)}
discount_rate <- o + n*g
# Build data frames for each scenario
T_00 <- data.frame(temp_increase = T_low,
                   year = t,
                   temp = temp(t, T_low)) %>% 
  mutate(economy = econ(temp)) %>% 
  mutate(consumption = consumption(economy, year)) %>% 
  mutate(utility = utility(consumption)) %>% 
  mutate(pv_utility = utility/(1+discount_rate)^year)

T_04.4 <- data.frame(temp_increase = T_bau,
                   year = t,
                   temp = temp(t, T_bau)) %>% 
  mutate(economy = econ(temp)) %>% 
  mutate(consumption = consumption(economy, year)) %>% 
  mutate(utility = utility(consumption)) %>% 
  mutate(pv_utility = utility/(1+discount_rate)^year) %>% 
  data.frame(T_00$pv_utility) %>% 
  mutate(pct_change = (pv_utility - T_00.pv_utility)/T_00.pv_utility*100)

L_o <- data.frame(sum(T_04.4$T_00.pv_utility),
                           sum(T_04.4$pv_utility)) %>%
  rename(pv_utility_base = sum.T_04.4.T_00.pv_utility.,
         pv_utility_bau = sum.T_04.4.pv_utility.) %>%
  mutate(L_o = (pv_utility_bau - pv_utility_base)/pv_utility_base) %>%
  mutate(L = L$L) %>%
  mutate(elasticity = .1/(L_o-L/L))
```


10% increase in n

```{r}
# Assign values
T_low <- 0
T_med <- 2
T_high <- 8
T_bau <- 4.4
t <- 0:200
o <- 0.005
n <- .5*1.1
g <- .01
b <- .05
# Build functions in this chunk
temp <- function(t, T) {temp = pmin(T*t/100, T)}
econ <- function(t) {econ = exp(-1*b*t^2)}
consumption <- function(k, t) {k*exp(g*t)}
utility <- function(c) {(c^(1-n))/(1-n)}
discount_rate <- o + n*g
# Build data frames for each scenario
T_00 <- data.frame(temp_increase = T_low,
                   year = t,
                   temp = temp(t, T_low)) %>% 
  mutate(economy = econ(temp)) %>% 
  mutate(consumption = consumption(economy, year)) %>% 
  mutate(utility = utility(consumption)) %>% 
  mutate(pv_utility = utility/(1+discount_rate)^year)

T_04.4 <- data.frame(temp_increase = T_bau,
                   year = t,
                   temp = temp(t, T_bau)) %>% 
  mutate(economy = econ(temp)) %>% 
  mutate(consumption = consumption(economy, year)) %>% 
  mutate(utility = utility(consumption)) %>% 
  mutate(pv_utility = utility/(1+discount_rate)^year) %>% 
  data.frame(T_00$pv_utility) %>% 
  mutate(pct_change = (pv_utility - T_00.pv_utility)/T_00.pv_utility*100)

L_n <- data.frame(sum(T_04.4$T_00.pv_utility),
                           sum(T_04.4$pv_utility)) %>%
  rename(pv_utility_base = sum.T_04.4.T_00.pv_utility.,
         pv_utility_bau = sum.T_04.4.pv_utility.) %>%
  mutate(L_n = (pv_utility_bau - pv_utility_base)/pv_utility_base) %>%
  mutate(L = L$L) %>%
  mutate(elasticity = .1/(L_n-L/L))
```

10% increase in g
```{r}
# Assign values
T_low <- 0
T_med <- 2
T_high <- 8
T_bau <- 4.4
t <- 0:200
o <- 0.005
n <- .5
g <- .01*1.1
b <- .05
# Build functions in this chunk
temp <- function(t, T) {temp = pmin(T*t/100, T)}
econ <- function(t) {econ = exp(-1*b*t^2)}
consumption <- function(k, t) {k*exp(g*t)}
utility <- function(c) {(c^(1-n))/(1-n)}
discount_rate <- o + n*g
# Build data frames for each scenario
T_00 <- data.frame(temp_increase = T_low,
                   year = t,
                   temp = temp(t, T_low)) %>% 
  mutate(economy = econ(temp)) %>% 
  mutate(consumption = consumption(economy, year)) %>% 
  mutate(utility = utility(consumption)) %>% 
  mutate(pv_utility = utility/(1+discount_rate)^year)

T_04.4 <- data.frame(temp_increase = T_bau,
                   year = t,
                   temp = temp(t, T_bau)) %>% 
  mutate(economy = econ(temp)) %>% 
  mutate(consumption = consumption(economy, year)) %>% 
  mutate(utility = utility(consumption)) %>% 
  mutate(pv_utility = utility/(1+discount_rate)^year) %>% 
  data.frame(T_00$pv_utility) %>% 
  mutate(pct_change = (pv_utility - T_00.pv_utility)/T_00.pv_utility*100)

L_g <- data.frame(sum(T_04.4$T_00.pv_utility),
                           sum(T_04.4$pv_utility)) %>%
  rename(pv_utility_base = sum.T_04.4.T_00.pv_utility.,
         pv_utility_bau = sum.T_04.4.pv_utility.) %>%
  mutate(L_g = (pv_utility_bau - pv_utility_base)/pv_utility_base) %>%
  mutate(L = L$L) %>%
  mutate(elasticity = .1/(L_g-L/L))
```



10% increase in b
```{r}
# Assign values
T_low <- 0
T_med <- 2
T_high <- 8
T_bau <- 4.4
t <- 0:200
o <- 0.005
n <- .5
g <- .01
b <- .05*1.1
# Build functions in this chunk
temp <- function(t, T) {temp = pmin(T*t/100, T)}
econ <- function(t) {econ = exp(-1*b*t^2)}
consumption <- function(k, t) {k*exp(g*t)}
utility <- function(c) {(c^(1-n))/(1-n)}
discount_rate <- o + n*g
# Build data frames for each scenario
T_00 <- data.frame(temp_increase = T_low,
                   year = t,
                   temp = temp(t, T_low)) %>% 
  mutate(economy = econ(temp)) %>% 
  mutate(consumption = consumption(economy, year)) %>% 
  mutate(utility = utility(consumption)) %>% 
  mutate(pv_utility = utility/(1+discount_rate)^year)

T_04.4 <- data.frame(temp_increase = T_bau,
                   year = t,
                   temp = temp(t, T_bau)) %>% 
  mutate(economy = econ(temp)) %>% 
  mutate(consumption = consumption(economy, year)) %>% 
  mutate(utility = utility(consumption)) %>% 
  mutate(pv_utility = utility/(1+discount_rate)^year) %>% 
  data.frame(T_00$pv_utility) %>% 
  mutate(pct_change = (pv_utility - T_00.pv_utility)/T_00.pv_utility*100)

L_b <- data.frame(sum(T_04.4$T_00.pv_utility),
                           sum(T_04.4$pv_utility)) %>%
  rename(pv_utility_base = sum.T_04.4.T_00.pv_utility.,
         pv_utility_bau = sum.T_04.4.pv_utility.) %>%
  mutate(L_b = (pv_utility_bau - pv_utility_base)/pv_utility_base) %>%
  mutate(L = L$L) %>%
  mutate(elasticity = .1/(L_b-L/L))
```


10% increase in T_bau
```{r}
# Assign values
T_low <- 0
T_med <- 2
T_high <- 8
T_bau <- 4.4*1.1
t <- 0:200
o <- 0.005
n <- .5
g <- .01
b <- .05
# Build functions in this chunk
temp <- function(t, T) {temp = pmin(T*t/100, T)}
econ <- function(t) {econ = exp(-1*b*t^2)}
consumption <- function(k, t) {k*exp(g*t)}
utility <- function(c) {(c^(1-n))/(1-n)}
discount_rate <- o + n*g
# Build data frames for each scenario
T_00 <- data.frame(temp_increase = T_low,
                   year = t,
                   temp = temp(t, T_low)) %>% 
  mutate(economy = econ(temp)) %>% 
  mutate(consumption = consumption(economy, year)) %>% 
  mutate(utility = utility(consumption)) %>% 
  mutate(pv_utility = utility/(1+discount_rate)^year)

T_04.4 <- data.frame(temp_increase = T_bau,
                   year = t,
                   temp = temp(t, T_bau)) %>% 
  mutate(economy = econ(temp)) %>% 
  mutate(consumption = consumption(economy, year)) %>% 
  mutate(utility = utility(consumption)) %>% 
  mutate(pv_utility = utility/(1+discount_rate)^year) %>% 
  data.frame(T_00$pv_utility) %>% 
  mutate(pct_change = (pv_utility - T_00.pv_utility)/T_00.pv_utility*100)

L_t <- data.frame(sum(T_04.4$T_00.pv_utility),
                           sum(T_04.4$pv_utility)) %>%
  rename(pv_utility_base = sum.T_04.4.T_00.pv_utility.,
         pv_utility_bau = sum.T_04.4.pv_utility.) %>%
  mutate(L_t = (pv_utility_bau - pv_utility_base)/pv_utility_base) %>%
  mutate(L = L$L) %>%
  mutate(elasticity = .1/(L_t-L/L))
```



2.c. 







